---
title: "2022: Day 1"
date: 2022-12-1
categories: [base R, lists]
draft: false
---

## Setup

[The original challenge](https://adventofcode.com/2022/day/1)

[My data](input){target="_blank"}

## Part 1

I'm including this post in the website example to demonstrate what a typical post will look like, using `post-template`, in the `_templates` directory as a starting point. This post is created by a call to `aoc_new_day(1, 2022)`. It can be deleted with a call to `aoc_delete_day(1, 2022)`, or all posts and the listing for the year can be deleted with `aoc_delete_year(2022)`.

```{r}
#| echo: false
OK <- "2022" < 3000
# Will only evaluate next code block if an actual year has been substituted for the placeholder
```

::: {.callout-note}
This is Ella Kaye's solution^[Ella is the author of this website template and of the **aochelpers** package, and the author of this demo post.], with her puzzle input. If attempting this challenge yourself, your solution will be different.
:::

```{r}
#| eval: !expr OK
library(aochelpers)
input <- aoc_input_vector(1, 2022, "numeric")
```

I'm using the `aoc_input_vector()` function from the **aochelpers** package to read in the data, but otherwise using base R functions (including the native pipe, `|>`) for this puzzle. 

In this challenge, we're given groups of numbers and we need to find the sum of each group. 
Our solution is the largest of these. The groups are separated by a blank line. When reading in the input as a numeric vector, these are coerced to `NA`.
We can identify the new groups by the `NA` values, produce an index for them with `cumsum(is.na(input))`,
which increments when a new `NA` is reached, then use this with `split()` to split the input into a list of vectors, one for each group.
We need the argument `na.rm = TRUE` in `sapply()` because each vector, other than the first, starts with `NA`, as that's where it was split.

```{r}
totals <- split(input, cumsum(is.na(input))) |> 
  sapply(sum, na.rm = TRUE) 

max(totals)
```

## Part 2

This is similar, except we want to find the sum of the sums of the top three groups.

```{r}
totals |> 
  sort() |> 
  tail(3) |> 
  sum()
```

##### Session info {.appendix}

<details><summary>Toggle</summary>

```{r}
#| echo: false
library(sessioninfo)
# save the session info as an object
pkg_session <- session_info(pkgs = "attached")

# get the quarto version
quarto_version <- system("quarto --version", intern = TRUE)

# inject the quarto info
pkg_session$platform$quarto <- paste(
  system("quarto --version", intern = TRUE), 
  "@", 
  quarto::quarto_path()
  )

# print it out
pkg_session
```

</details>





